---
title: "AIC/BIC Logistic Regression Health"
author: "Michelle Le 30145965"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(dplyr) # data manipulation
library(naniar) # visualizing non-response
library(VIM) # visualizing non-response
library(mice) # multiple imputation
library(naniar) # summarize missingness patterns
library(ggplot2) # general data visualization
library(patchwork) # formatting data visualization
library(poLCA) # lca modelling
library(survey)
library(car)
```

# Loading the Data

```{r load data}
# data <- read.csv(file.choose(), colClasses = "character") # run within R chunk
data <- read.csv("/Users/michellele/Downloads/data 501/5012-NGS-END_2020-CSV/ngs2020_pumf.csv", colClasses = "character") # run to knit as pdf
# head(data)

# check the number of rows
nrow(data)

# filter the data to bachelors only
data <- data %>%
  filter(CERTLEVP == 2) # CERTLEVP refers to university certificate or diploma below a bachelor's degree AND bachelor's degree

nrow(data)
```


# Preliminary Data Cleaning

```{r prelim cleaning}
# check how many columns are in the dataset
ncol_before <- ncol(data)
ncol_before

# we remove the following columns because they define employment outcomes and are not graduate characteristics
# although we do keep LFSTATP for later***
# LFSTATP is employment status (employed, unemployed, etc.)
oth_cols_to_remove <- c("CERTLEVP", # remove since we're only looking at bachelors
                        "LMA_010",
                        "LMA2_07",
                        "LMA3_P01",
                        # "LFCINDP", # keep since we will later need for job-sector specific analysis
                        "LFCOCCP",
                        "LFWFTPTP",
                        "LMA6_05",
                        "LMA6_08",
                        "JOBQLEVP",
                        "JOBQLGRD",
                        "JOBQLINT",
                        "LMA6_11",
                        "LMA6_12",
                        "LMA6_13A",
                        "LMA6_13B",
                        "LMA6_13C",
                        "JOBINCP",
                        "LMA6_15",
                        "AFT_P010",
                        "AFT_P020",
                        "AFT_P040",
                        "AFT_050",
                        "AFT_070",
                        "AFT_080",
                        "AFT_090",
                        "WTPF", # remove the weights for now so they don't get accidentally imputed; although will need them later
                        "VERDATE",
                        "PERSINCP")

data <- data[ , !(names(data) %in% oth_cols_to_remove)]

# check how many columns are left
ncol_after <- ncol(data)
ncol_after

# some columns use "99" to denote not-stated whereas other columns use "9"; we need to distinguish between these columns

# colnames(data)

cols_99 <- c("PGMCIPAP", 
             "PGM_410",
             "PGM_415",
             "STL_030",
             "BEF_160",
             "LFCINDP"
             )
length(cols_99)
cols_9 <- setdiff(names(data), c(cols_99, "PUMFID", "LFSTATP")) # every other column uses "9", do not include "PUMFID", "LFSTATP"
cols_9
length(cols_9)

# replace "not-stated" (denoted by either "99" or "9") with NA
data <- data %>%
  mutate(
    across(all_of(cols_99), ~na_if(., "99")),
    across(all_of(cols_9),  ~na_if(., "9"))
  )
dim(data)
```

Moving forward, we will only be working with data applicable to all respondents. Whether a question was applicable to all respondents is defined by the universe. The universe for each question can be found in the `NGS2020-PUMF_E.pdf` document.  The following variables were applicable to all respondents:

```{r get relevant columns}
allresp_universe <- c(# "CERTLEVP", # program - level of study
                      "REG_INST", # region of inst
                      "REG_RESP", # region of res 2023
                      "PGMCIPAP", # 2020 program
                      "PGM_P034", # 2020 full-time vs part-time
                      "PGM_P100", # work placement during program status
                      "PGM_280A", # started a business
                      "PGM_280B", # completed entrepreneurship courses
                      "PGM_280C", # business plan or pitch comp
                      "PGM_280D", # visited entrepreneurship centre
                      "PGM_280E", # worked on entrepreneurship project
                      "PGM_280F", # entrepreneurship - none of the above
                      "PGM_290", # worked during the program
                      "PGM_350", # volunteered during the program
                      "PGM_380", # completed components taken outside of canada
                      "PGM_P401", # online/distance education prior to COVID-19
                      "PGM_410", # main factor in choice of inst
                      "PGM_415", # main factor in choice of program
                      "PGM_430", # choose to repeat field of study
                      "EDU_010", # other programs taken since 2020
                      "HLOSINTP", # 2023 highest level of educ comp
                      "STL_010", # applied for student loan
                      "STULOANS", # received student loan
                      "STL_160B", # RESP for funding
                      "STL_160C", # grant/bursary for funding
                      "STL_160D", # non-gov grant/bursary for funding
                      "STL_160E", # scholarship/award for funding
                      "STL_160F", # employment earnings/savings for fund
                      "STL_160G", # research/TA for funding
                      "STL_160H", # parents, family, friends for funding
                      "STL_160I", # bank loan for funding
                      "STL_160J", # credit card for funding
                      "STL_160L", # employer for funding
                      "STLP160N", # other for funding
                      # "DBTALGRD", # debt; universe != all respondents
                      "BEF_P140", # main activity before 2020 program
                      "BEF_160", # work exp before 2020 program
                      "HLOSGRDP", # 2020 highest level of educ comp
                      "PAR1GRD", # parent 1 education 2020
                      "PAR1INT", # parent 1 education 2023
                      "PAR2GRD", # parent 2 education 2020
                      "PAR2INT", # parent 2 education 2023
                      "GRADAGEP", # 2020 age
                      "GENDER2", # gender
                      "MS_P01", # curr marital status
                      "MS_P02", # dependent children status
                      "CTZSHIPP", # 2023 citizenship status
                      "VISBMINP", # visible minority membership
                      "DDIS_FL" # disability status
                      )

# drop columns not applicable to all respondents
data_reduced <- data[ , (names(data) %in% allresp_universe)]
data_reduced$LFSTATP <- data$LFSTATP # add back LFSTATP
data_reduced$PUMFID <- data$PUMFID # add back PUMFID
data_reduced$LFCINDP <- data$LFCINDP # add back LFCINDP
dim(data_reduced)
```

# Check Missing Data

```{r check missing data}
# check whether the response has any missing data
sum(is.na(data_reduced$LFSTATP))

# check whether the ID column has any missing data
sum(is.na(data_reduced$PUMFID))

# check whether the LFCINDP column has any missing data
sum(is.na(data_reduced$LFCINDP))

# check the number of non-responses per column
survey_cols <- intersect(c(cols_99, cols_9), names(data_reduced))
col_missing <- colSums(is.na(data_reduced[, survey_cols]))
sort(col_missing, decreasing = TRUE)
top_missing_cols <- names(sort(col_missing, decreasing = TRUE)[1:5])

# check which column has the most non-responses
which.max(colSums(is.na(data_reduced[, survey_cols]))) # output is the column name and the corresponding index

# check the non-response rate per column
missing_prop <- colMeans(is.na(data_reduced[, survey_cols]))
round(missing_prop*100, 2) # multiply by 100 to get a percentage; easier to read

# check the number of non-responses per respondent
row_missing <- rowSums(is.na(data_reduced[, survey_cols]))
# row_missing

# check the proportion of non-responses per respondent
row_missing_prop <- row_missing / length(survey_cols)
# row_missing_prop

respondent_nonresp <- data.frame(row_missing, row_missing_prop)
summary(row_missing)
summary(row_missing_prop)

# visual summary
gg_miss_upset(data_reduced[, survey_cols])  # which combinations of NAs occur
```

The above visualization summarizes the most common combinations of survey questions left unanswered. Notably, parent's education level (`PAR2GRD`, `PAR2INT`, `PAR1GRD`, `PAR1INT`) and citizenship status (`CTZSHIPP`) were most commonly unanswered.

```{r examining distribution of missing data}

p1 <- ggplot(data_reduced, aes(x = PAR2GRD)) +
  geom_bar() +
  labs(title = "Distribution of PAR2GRD (Before Imputation)",
       x = "PAR2GRD",
       y = "Count")

p2 <- ggplot(data_reduced, aes(x = PAR2INT)) +
  geom_bar() +
  labs(title = "Distribution of PAR2INT (Before Imputation)",
       x = "PAR2INT",
       y = "Count")

p3 <- ggplot(data_reduced, aes(x = CTZSHIPP)) +
  geom_bar() +
  labs(title = "Distribution of CTZSHIPP (Before Imputation)",
       x = "CTZSHIPP",
       y = "Count")

p4 <- ggplot(data_reduced, aes(x = PAR1GRD)) +
  geom_bar() +
  labs(title = "Distribution of PAR1GRD (Before Imputation)",
       x = "PAR1GRD",
       y = "Count")

p5 <- ggplot(data_reduced, aes(x = PAR1INT)) +
  geom_bar() +
  labs(title = "Distribution of PAR1INT (Before Imputation)",
       x = "PAR1INT",
       y = "Count")

p1
p2
p3
p4
p5
```

# Imputation

We will use Multiple Imputation with Chained Equations (MICE) to handle the missing data.

```{r imputation}
# exclude PUMFID and LFSTATP since they are not missing
data_impute <- data_reduced %>%
  dplyr::select(-PUMFID)
data_impute <- data_impute %>%
  dplyr::select(-LFSTATP)

names(data_impute)

# convert to factor
data_impute <- data_impute %>%
  mutate(across(everything(), as.factor))

# imp <- mice(data_impute,
#             m = 5, # create 5 datasets
#             maxit = 20, # run 20 iterations for stability
#             method = "cart", # use cart
#             seed = 123) # use seed for reproducability
# 
# # diagnostic plots to ensure good mixing
# # plot(imp)
# # stripplot(imp, pch = 20, cex = 1.2)
# # summary(imp)
# 
# # for now, we'll take the first complete dataset out of the 5 generated and use this for our lca
# data_completed <- complete(imp, 1)

# # write each of the five datasets to a csv so we don't have to run mice again
# # for later analysis, we will need to pool the 5 datasets(?)
# for (i in 1:5) {
#   # get the ith completed dataset
#   completed_data <- complete(imp, i)
#   
#   # create a file name
#   file_name <- paste0("data_completed_imp", i, ".csv")
#   
#   # write to CSV
#   write.csv(completed_data, file_name, row.names = FALSE)
# }

data_completed <- read.csv("/Users/michellele/Downloads/data 501/data_completed_imp1.csv", colClasses = "character")

# add back the ID and employment status columns
data_completed$PUMFID <- data_reduced$PUMFID
data_completed$LFSTATP <- data_reduced$LFSTATP


# check the distributions
p6 <- ggplot(data_completed, aes(x = PAR2GRD)) +
  geom_bar() +
  labs(title = "Distribution of PAR2GRD (After Imputation)",
       x = "PAR2GRD",
       y = "Count")

p7 <- ggplot(data_completed, aes(x = PAR2INT)) +
  geom_bar() +
  labs(title = "Distribution of PAR2INT (After Imputation)",
       x = "PAR2INT",
       y = "Count")

p8 <- ggplot(data_completed, aes(x = CTZSHIPP)) +
  geom_bar() +
  labs(title = "Distribution of CTZSHIPP (After Imputation)",
       x = "CTZSHIPP",
       y = "Count")

p9 <- ggplot(data_completed, aes(x = PAR1GRD)) +
  geom_bar() +
  labs(title = "Distribution of PAR1GRD (After Imputation)",
       x = "PAR1GRD",
       y = "Count")

p10 <- ggplot(data_completed, aes(x = PAR1INT)) +
  geom_bar() +
  labs(title = "Distribution of PAR1INT (After Imputation)",
       x = "PAR1INT",
       y = "Count")

p1 + p6
p2 + p7
p3 + p8
p4 + p9
p5 + p10
```

From these plots, we can observe that the distribution before and after MICE is relatively similar.

# LCA Indicator Selection

With LCA, it is not the case that more indicators is better. Too many indicators will increase model complexity, reduce interpretability and cause convergence problems. Moreover, good LCA models should produce clear separation between classes. Thus, indicators that are composite of other indicators should be excluded.

Therefore, to guide our feature selection, we need to consider a conceptual framework, exclude variables that break LCA assumptions such as conditional independence, and exclude low-information variables (i.e., small variance).

While factors such as the region of the respondent's institution and resident may provide some useful insight, we are not particularly interested in these indicators. Similarly, we are not interested in the respondent's main factor in their choice of institution and/or program or whether they would choose the same program again if they could. That is, regional and institutional variables are not part of our conceptual framework.

To some extent, we are interested in the respondent's entrepreneurship skills. Particularly, whether or not they were trained in or had exposure to them, not how. For this reason, we will collapse the entrepreneurship skills section into one variable.

```{r}
data_completed$entrepreneurship_any <- ifelse(
  data_completed$PGM_280A == 1 |
  data_completed$PGM_280B == 1 |
  data_completed$PGM_280C == 1 |
  data_completed$PGM_280D == 1 |
  data_completed$PGM_280E == 1,
  1, 2
)
```


Similarly, we are somewhat interested in how the respondent financed their education (specifically, whether or not they received student loans).

Based on this criteria we will tentatively explore:

* `PGMCIPAP`
* `PGM_P034`
* `PGM_P100`
* `entrepreneurship_any`
* `PGM_290`
* `PGM_350`
* `PGM_380`
* `PGM_P401`
* `EDU_010`
* `HLOSINTP`
* `STULOANS`
* `BEF_P140`
* `BEF_160`
* `PAR1GRD`
* `PAR1INT`
* `PAR2GRD`
* `PAR2INT`
* `GRADAGEP`
* `GENDER2`
* `CTZSHIPP`
* `VISBMINP`
* `DDIS_FL`

We can also narrow these down further by evaluating their variances. We will not move forward with any variables that present limited variance.

```{r examining variance}
potential_lca_vars <- c(
  "PGMCIPAP", # 2020 Program - Aggregated CIP 2021
  "PGM_P034", # 2020 Program - Full-time or part-time student
  "PGM_P100", # Work placement during program
  "entrepreneurship_any",
  "PGM_290", # 2020 Program - Worked during program
  "PGM_350", # 2020 Program - Volunteer activities during program
  "PGM_380", # 2020 Program - Components taken outside of Canada
  "PGM_P401", # 2020 Program - Online or distance education
  "EDU_010", # After 2020 program - Other postsecondary programs taken
  "STULOANS", # Government-student loan program - Received
  "BEF_P140", # Before 2020 Program - Main activity during 12 months before
  "BEF_160", # Before 2020 program - Number of months of work experience
  "PAR1GRD", # 2020 Program - Level of education vs of the other parent
  "PAR1INT", # 2023 Program - Level of education vs of the other parent
  "PAR2GRD", # 2020 Program - Level of education vs of the other parent
  "PAR2INT", # 2023 Program - Level of education vs of the other parent
  "GRADAGEP", # 2020 Program - Age at time of graduation - Grouping
  "GENDER2", # Gender after distribution of non-binary persons
  "CTZSHIPP", # Time of interview 2023 - Status in Canada
  "VISBMINP", # Self-identified as a member of a visible minority group
  "DDIS_FL" # Disability status
)

for(v in potential_lca_vars){
  
  tab <- table(data_completed[[v]])
  
  pie(tab,
      main = paste("Distribution of", v),
      labels = paste0(names(tab), " (", round(prop.table(tab)*100,1), "%)"))
}

# compute the largest proportions for each column
max_prop <- sapply(data_completed[potential_lca_vars], function(x) {
  max(prop.table(table(x)))
})

# remove them from the lca analysis
lca_vars <- potential_lca_vars[max_prop <= 0.95]

removed_vars <- potential_lca_vars[max_prop > 0.95]
removed_vars

# is there really a difference between `PAR1GRD` and `PAR1INT`/`PAR2GRD` and `PAR2INT`?
# we consider dropping the latter from our analysis as they may be composite features
lca_vars <- lca_vars[!lca_vars %in% c("PAR1INT", "PAR2INT")]

# check how many indicators there are
length(lca_vars)
```

# LCA

```{r lca analysis,  eval=FALSE}
# subset the data
lca_data <- data_completed %>%
  dplyr::select("PUMFID", "LFSTATP", dplyr::all_of(lca_vars))

# convert LCA indicators to factors if not already
lca_data[lca_vars] <- lapply(lca_data[lca_vars], as.factor)

# define the formula
f <- as.formula(
  paste("cbind(", paste(lca_vars, collapse = ", "), ") ~ 1")
)

# run with 2-8 classes
lca_2 <- poLCA(f, lca_data, nclass = 2, maxiter = 5000, nrep = 10)
lca_3 <- poLCA(f, lca_data, nclass = 3, maxiter = 5000, nrep = 10)
lca_4 <- poLCA(f, lca_data, nclass = 4, maxiter = 5000, nrep = 10)
lca_5 <- poLCA(f, lca_data, nclass = 5, maxiter = 5000, nrep = 10)
lca_6 <- poLCA(f, lca_data, nclass = 6, maxiter = 5000, nrep = 10)
lca_7 <- poLCA(f, lca_data, nclass = 7, maxiter = 5000, nrep = 10)
lca_8 <- poLCA(f, lca_data, nclass = 8, maxiter = 5000, nrep = 10)

N <- nrow(lca_data)

sabic_values <- c(
  -2 * lca_2$llik + log((N + 2)/24) * lca_2$npar,
  -2 * lca_3$llik + log((N + 2)/24) * lca_3$npar,
  -2 * lca_4$llik + log((N + 2)/24) * lca_4$npar,
  -2 * lca_5$llik + log((N + 2)/24) * lca_5$npar,
  -2 * lca_6$llik + log((N + 2)/24) * lca_6$npar,
  -2 * lca_7$llik + log((N + 2)/24) * lca_7$npar,
  -2 * lca_8$llik + log((N + 2)/24) * lca_8$npar
)

# compare model fit using aic, bic, and sabic
data.frame(
  Classes = 2:8,
  AIC = c(lca_2$aic, lca_3$aic, lca_4$aic, lca_5$aic, lca_6$aic, lca_7$aic, lca_8$aic),
  BIC = c(lca_2$bic, lca_3$bic, lca_4$bic, lca_5$bic, lca_6$bic, lca_7$bic, lca_8$bic),
  SABIC = sabic_values
)

# also take a look at entropy
compute_entropy <- function(model) {
  posterior <- model$posterior
  K <- ncol(posterior)
  N <- nrow(posterior)
  
  ent <- -sum(ifelse(posterior > 0, posterior * log(posterior), 0)) / N
  ent <- ent / log(K)
  1 - ent
}

entropy_2 <- compute_entropy(lca_2)
entropy_3 <- compute_entropy(lca_3)
entropy_4 <- compute_entropy(lca_4)
entropy_5 <- compute_entropy(lca_5)
entropy_6 <- compute_entropy(lca_6)
entropy_7 <- compute_entropy(lca_7)
entropy_8 <- compute_entropy(lca_8)

entropy_values <- c(entropy_2, entropy_3, entropy_4, entropy_5, entropy_6, entropy_7, entropy_8)
entropy_values

# check class sizes
prop.table(table(lca_2$predclass))
prop.table(table(lca_3$predclass))
prop.table(table(lca_4$predclass))
prop.table(table(lca_5$predclass))
prop.table(table(lca_6$predclass))
prop.table(table(lca_7$predclass))
prop.table(table(lca_8$predclass))

models <- list(
  lca_2,
  lca_3,
  lca_4,
  lca_5,
  lca_6,
  lca_7,
  lca_8
)

names(models) <- 2:8

# Define the pseudo-LR test function
LR_test <- function(model_k, model_k_minus1) {
  LR_stat <- -2 * (model_k_minus1$llik - model_k$llik)
  df <- model_k$npar - model_k_minus1$npar
  p_value <- 1 - pchisq(LR_stat, df)
  c(LR_stat = LR_stat, df = df, p_value = p_value)
}

# Initialize a data frame to store LR results
LR_results <- data.frame(
  Classes = 3:8,   # start at 3 because we compare to k-1
  LR_Stat = NA,
  LR_df = NA,
  LR_p = NA
)

# Loop through k=3 to 8 and compare with k-1
for (k in 3:8) {
  LR <- LR_test(models[[as.character(k)]], models[[as.character(k-1)]])
  LR_results[LR_results$Classes == k, c("LR_Stat", "LR_df", "LR_p")] <- LR
}

LR_results
lca_data$class <- lca_5$predclass
lca_5
```

Based on the above results, the model with 5 classes is most optimal. In general, the differences in AIC, BIC, SABIC, and entropy across classes are negligible. 5 classes balances low AIC, BIC, and SABIC with high entropy, and moderate class size proportions.

```{r}
# lca_data$class <- lca_5$predclass
# lca_5
```
```{r load other imputed datasets}
data_completed_1 <- read.csv("/Users/michellele/Downloads/data 501/data_completed_imp1.csv", colClasses = "character")
data_completed_1$PUMFID <- data_reduced$PUMFID
data_completed_1$LFSTATP <- data_reduced$LFSTATP

data_completed_2 <- read.csv("/Users/michellele/Downloads/data 501/data_completed_imp2.csv", colClasses = "character")
data_completed_2$PUMFID <- data_reduced$PUMFID
data_completed_2$LFSTATP <- data_reduced$LFSTATP

data_completed_3 <- read.csv("/Users/michellele/Downloads/data 501/data_completed_imp3.csv", colClasses = "character")
data_completed_3$PUMFID <- data_reduced$PUMFID
data_completed_3$LFSTATP <- data_reduced$LFSTATP

data_completed_4 <- read.csv("/Users/michellele/Downloads/data 501/data_completed_imp4.csv", colClasses = "character")
data_completed_4$PUMFID <- data_reduced$PUMFID
data_completed_4$LFSTATP <- data_reduced$LFSTATP

data_completed_5 <- read.csv("/Users/michellele/Downloads/data 501/data_completed_imp5.csv", colClasses = "character")
data_completed_5$PUMFID <- data_reduced$PUMFID
data_completed_5$LFSTATP <- data_reduced$LFSTATP

# str(data_completed_1)
# str(data_completed_2)
# str(data_completed_3)
# str(data_completed_4)
# str(data_completed_5)

data_completed_1[] <- lapply(data_completed_1, function(x) {
  if (is.character(x)) as.factor(x) else x
})
# str(data_completed_1)

data_completed_2[] <- lapply(data_completed_2, function(x) {
  if (is.character(x)) as.factor(x) else x
})
# str(data_completed_2)

data_completed_3[] <- lapply(data_completed_3, function(x) {
  if (is.character(x)) as.factor(x) else x
})
# str(data_completed_3)

data_completed_4[] <- lapply(data_completed_4, function(x) {
  if (is.character(x)) as.factor(x) else x
})
# str(data_completed_4)

data_completed_5[] <- lapply(data_completed_5, function(x) {
  if (is.character(x)) as.factor(x) else x
})
# str(data_completed_5)

# filter out rows where LFSTATP != 1, 2
data_filtered_1 <- data_completed_1[data_completed_1$LFSTATP %in% c(1, 2), ]
data_filtered_2 <- data_completed_2[data_completed_2$LFSTATP %in% c(1, 2), ]
data_filtered_3 <- data_completed_3[data_completed_3$LFSTATP %in% c(1, 2), ]
data_filtered_4 <- data_completed_4[data_completed_4$LFSTATP %in% c(1, 2), ]
data_filtered_5 <- data_completed_5[data_completed_5$LFSTATP %in% c(1, 2), ]

# Filtering for program 9
data_filtered_1 = filter(data_filtered_1, PGMCIPAP == 9)
data_filtered_2 = filter(data_filtered_2, PGMCIPAP == 9)
data_filtered_3 = filter(data_filtered_3, PGMCIPAP == 9)
data_filtered_4 = filter(data_filtered_4, PGMCIPAP == 9)
data_filtered_5 = filter(data_filtered_5, PGMCIPAP == 9)

dim(data_filtered_1)
dim(data_filtered_2)
dim(data_filtered_3)
dim(data_filtered_4)
dim(data_filtered_5)

data_filtered_1$LFSTATP = ifelse(data_filtered_1$LFSTATP == 2,0,1) # unemployed = 0; employed = 1
data_filtered_2$LFSTATP = ifelse(data_filtered_2$LFSTATP == 2,0,1)
data_filtered_3$LFSTATP = ifelse(data_filtered_3$LFSTATP == 2,0,1)
data_filtered_4$LFSTATP = ifelse(data_filtered_4$LFSTATP == 2,0,1)
data_filtered_5$LFSTATP = ifelse(data_filtered_5$LFSTATP == 2,0,1)
```

# Weights set-up
This section adds back the weights too their corresponding imputed dataset based on PUMFID
```{r}

data_full <- read.csv("/Users/michellele/Downloads/data 501/5012-NGS-END_2020-CSV/ngs2020_pumf.csv", colClasses = "character")

data_filtered_1 <- data_filtered_1 %>%
  left_join(
    dplyr::select(data_full, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join lca_data$PUMFID to data$PUMFID
  )

data_filtered_2 <- data_filtered_2 %>%
  left_join(
    dplyr::select(data_full, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join lca_data$PUMFID to data$PUMFID
  )

data_filtered_3 <- data_filtered_3 %>%
  left_join(
    dplyr::select(data_full, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join lca_data$PUMFID to data$PUMFID
  )

data_filtered_4 <- data_filtered_4 %>%
  left_join(
    dplyr::select(data_full, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join lca_data$PUMFID to data$PUMFID
  )

data_filtered_5 <- data_filtered_5 %>%
  left_join(
    dplyr::select(data_full, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join lca_data$PUMFID to data$PUMFID
  )

# Converting WTPF as numeric
data_filtered_1$WTPF = as.numeric(data_filtered_1$WTPF)
data_filtered_2$WTPF = as.numeric(data_filtered_2$WTPF) 
data_filtered_3$WTPF = as.numeric(data_filtered_3$WTPF)
data_filtered_4$WTPF = as.numeric(data_filtered_4$WTPF)
data_filtered_5$WTPF = as.numeric(data_filtered_5$WTPF)
```

Creating training (70%) and test (30%) sets
```{r}
# Creating training and test sets

set.seed(1)
train <- sample(nrow(data_filtered_1), 0.7*nrow(data_filtered_1), replace = FALSE)

train_1 <- data_filtered_1[train,]
train_2 <- data_filtered_2[train,]
train_3 <- data_filtered_3[train,]
train_4 <- data_filtered_4[train,]
train_5 <- data_filtered_5[train,]


test_1 <- data_filtered_1[-train,]
test_2 <- data_filtered_2[-train,]
test_3 <- data_filtered_3[-train,]
test_4 <- data_filtered_4[-train,]
test_5 <- data_filtered_5[-train,]
```

```{r prepare for stepwise AIC/BIC logistic regression modeling}
predictors_1 <- setdiff(names(data_filtered_1), c("LFCINDP", "PUMFID", "LFSTATP", "WTPF"))
predictors_2 <- setdiff(names(data_filtered_2), c("LFCINDP", "PUMFID", "LFSTATP", "WTPF"))
predictors_3 <- setdiff(names(data_filtered_3), c("LFCINDP", "PUMFID", "LFSTATP", "WTPF"))
predictors_4 <- setdiff(names(data_filtered_4), c("LFCINDP", "PUMFID", "LFSTATP", "WTPF"))
predictors_5 <- setdiff(names(data_filtered_5), c("LFCINDP", "PUMFID", "LFSTATP", "WTPF"))

formula_1 <- as.formula(
  paste("LFSTATP ~", paste(predictors_1, collapse = " + "))
)
formula_2 <- as.formula(
  paste("LFSTATP ~", paste(predictors_2, collapse = " + "))
)
formula_3 <- as.formula(
  paste("LFSTATP ~", paste(predictors_3, collapse = " + "))
)
formula_4 <- as.formula(
  paste("LFSTATP ~", paste(predictors_4, collapse = " + "))
)
formula_5 <- as.formula(
  paste("LFSTATP ~", paste(predictors_5, collapse = " + "))
)

full_model_1 <- glm(formula_1,
             data = data_filtered_1, subset = train,
             family = binomial)
full_model_2 <- glm(formula_2,
             data = data_filtered_2, subset = train,
             family = binomial)
full_model_3 <- glm(formula_3,
             data = data_filtered_3, subset = train,
             family = binomial)
full_model_4 <- glm(formula_4,
             data = data_filtered_4, subset = train,
             family = binomial)
full_model_5 <- glm(formula_5,
             data = data_filtered_5, subset = train,
             family = binomial)
```

```{r stepwise aic logistic regression}
aic_step_model_1 <- step(full_model_1, direction = "both", trace = 0)
# summary(aic_step_model_1)

aic_step_model_2 <- step(full_model_2, direction = "both", trace = 0)
# summary(aic_step_model_2)

aic_step_model_3 <- step(full_model_3, direction = "both", trace = 0)
# summary(aic_step_model_3)

aic_step_model_4 <- step(full_model_4, direction = "both", trace = 0)
# summary(aic_step_model_4)

aic_step_model_5 <- step(full_model_5, direction = "both", trace = 0)
# summary(aic_step_model_5)
```

# Check Assumptions for AIC Model
```{r}
# Assumption 1: No severe multicollinearity
vif(aic_step_model_1)
vif(aic_step_model_2)
vif(aic_step_model_3)
vif(aic_step_model_4)
vif(aic_step_model_5)

# Assumption 2: Checking for influential points using Cook's Distance
cooks_1 <- which(cooks.distance(aic_step_model_1) > 4/nrow(data_filtered_1[train,]))
# cooks_1
cooks_2 <- which(cooks.distance(aic_step_model_2) > 4/nrow(data_filtered_2[train,]))
# cooks_2
cooks_3 <- which(cooks.distance(aic_step_model_3) > 4/nrow(data_filtered_3[train,]))
# cooks_3
cooks_4 <- which(cooks.distance(aic_step_model_4) > 4/nrow(data_filtered_4[train,]))
# cooks_4
cooks_5 <- which(cooks.distance(aic_step_model_5) > 4/nrow(data_filtered_5[train,]))
# cooks_5
```

```{r stepwise bic logistic regression}
bic_step_model_1 <- step(full_model_1, k = log(nrow(data_filtered_1)), trace = 0)
# summary(bic_step_model_1)

bic_step_model_2 <- step(full_model_2, k = log(nrow(data_filtered_2)), trace = 0)
# summary(bic_step_model_2)

bic_step_model_3 <- step(full_model_3, k = log(nrow(data_filtered_3)), trace = 0)
# summary(bic_step_model_3)

bic_step_model_4 <- step(full_model_4, k = log(nrow(data_filtered_4)), trace = 0)
# summary(bic_step_model_4)

bic_step_model_5 <- step(full_model_5, k = log(nrow(data_filtered_5)), trace = 0)
# summary(bic_step_model_5)
```

# Check Assumptions for BIC Model
```{r}
# Assumption 1: No severe multicollinearity
vif(bic_step_model_1)
vif(bic_step_model_2)
vif(bic_step_model_3)
vif(bic_step_model_4)
vif(bic_step_model_5)

# Assumption 2: Checking for influential points using Cook's Distance
cooks_1 <- which(cooks.distance(bic_step_model_1) > 4/nrow(data_filtered_1[train,]))
# cooks_1
cooks_2 <- which(cooks.distance(bic_step_model_2) > 4/nrow(data_filtered_2[train,]))
# cooks_2
cooks_3 <- which(cooks.distance(bic_step_model_3) > 4/nrow(data_filtered_3[train,]))
# cooks_3
cooks_4 <- which(cooks.distance(bic_step_model_4) > 4/nrow(data_filtered_4[train,]))
# cooks_4
cooks_5 <- which(cooks.distance(bic_step_model_5) > 4/nrow(data_filtered_5[train,]))
# cooks_5
```

```{r check common aic predictors}
aic_vars_1 <- setdiff(names(coef(aic_step_model_1)), "(Intercept)")
aic_vars_2 <- setdiff(names(coef(aic_step_model_2)), "(Intercept)")
aic_vars_3 <- setdiff(names(coef(aic_step_model_3)), "(Intercept)")
aic_vars_4 <- setdiff(names(coef(aic_step_model_4)), "(Intercept)")
aic_vars_5 <- setdiff(names(coef(aic_step_model_5)), "(Intercept)")

all_aic_vars <- c(aic_vars_1, aic_vars_2, aic_vars_3, aic_vars_4, aic_vars_5)
aic_tab <- table(all_aic_vars)
common_aic_vars <- names(aic_tab[aic_tab == 5])
common_aic_vars
```

```{r check common bic predictors}
bic_vars_1 <- setdiff(names(coef(bic_step_model_1)), "(Intercept)")
bic_vars_2 <- setdiff(names(coef(bic_step_model_2)), "(Intercept)")
bic_vars_3 <- setdiff(names(coef(bic_step_model_3)), "(Intercept)")
bic_vars_4 <- setdiff(names(coef(bic_step_model_4)), "(Intercept)")
bic_vars_5 <- setdiff(names(coef(bic_step_model_5)), "(Intercept)")

all_bic_vars <- c(bic_vars_1, bic_vars_2, bic_vars_3, bic_vars_4, bic_vars_5)
bic_tab <- table(all_bic_vars)
common_bic_vars <- names(bic_tab[bic_tab == 5])
common_bic_vars
```


