---
title: "Data_501_LogisticRegression"
author: "Juliana Los Banos"
date: "2026-03-15"
output: html_document
---

#Libraries needed
```{r}
library(dplyr) # data manipulation
library(mice) #for Rubin's rule
library(car) #for VIF
library(pROC) #for predicting and roc curve
library(caret) #roc curve
library(Metrics) #for performance metrics
library(survey) #for survey weights application

```

#Loading data
```{r load data}

data <- read.csv("C:/Users/janal/OneDrive/Desktop/DATA 501/ngs2020_pumf.csv", colClasses = "character") # run to knit as pdf

# filter the data to bachelors only
data <- data %>%
  filter(CERTLEVP == 2) # CERTLEVP refers to university certificate or diploma below a bachelor's degree AND bachelor's degree

#View(data)
```


#Loading imputed data
```{r}

#read all the imputed datasets
imp_data1 <- read.csv("C:/Users/janal/OneDrive/Desktop/DATA 501/data_completed_imp1.csv", colClasses = "character")
imp_data2 <- read.csv("C:/Users/janal/OneDrive/Desktop/DATA 501/data_completed_imp2.csv", colClasses = "character")
imp_data3 <- read.csv("C:/Users/janal/OneDrive/Desktop/DATA 501/data_completed_imp3.csv", colClasses = "character")
imp_data4 <- read.csv("C:/Users/janal/OneDrive/Desktop/DATA 501/data_completed_imp4.csv", colClasses = "character")
imp_data5 <- read.csv("C:/Users/janal/OneDrive/Desktop/DATA 501/data_completed_imp5.csv", colClasses = "character")

# add back the ID (PUMFID) and employment status (LFSTATP) columns
imp_data1$PUMFID <- data$PUMFID
imp_data1$LFSTATP <- data$LFSTATP

imp_data2$PUMFID <- data$PUMFID
imp_data2$LFSTATP <- data$LFSTATP

imp_data3$PUMFID <- data$PUMFID
imp_data3$LFSTATP <- data$LFSTATP

imp_data4$PUMFID <- data$PUMFID
imp_data4$LFSTATP <- data$LFSTATP

imp_data5$PUMFID <- data$PUMFID
imp_data5$LFSTATP <- data$LFSTATP


```

#Weights set-up

This section adds back the survey weights to their corresponding imputed dataset based on PUMFID
```{r}


imp_data1 <- imp_data1 %>%
  left_join(
    dplyr::select(data, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join imp_data1$PUMFID to data$PUMFID
  )

imp_data2 <- imp_data2 %>%
  left_join(
    dplyr::select(data, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join imp_data2$PUMFID to data$PUMFID
  )

imp_data3 <- imp_data3 %>%
  left_join(
    dplyr::select(data, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join imp_data3$PUMFID to data$PUMFID
  )

imp_data4 <- imp_data4 %>%
  left_join(
    dplyr::select(data, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join imp_data4$PUMFID to data$PUMFID
  )

imp_data5 <- imp_data5 %>%
  left_join(
    dplyr::select(data, PUMFID, WTPF),  # select columns from `data`
    by = c("PUMFID" = "PUMFID")          # join imp_data5$PUMFID to data$PUMFID
  )

#Converting WTPF (weight variable) as numeric
imp_data1$WTPF = as.numeric(imp_data1$WTPF)
imp_data2$WTPF = as.numeric(imp_data2$WTPF) 
imp_data3$WTPF = as.numeric(imp_data3$WTPF)
imp_data4$WTPF = as.numeric(imp_data4$WTPF)
imp_data5$WTPF = as.numeric(imp_data5$WTPF)
```


#Data preprocessing
Preparing imputed data for logistic regression (Field of study 9: Health and other related fields)
```{r}

#Filtering for only Employed versus Unemployed labour force status
imp_data1 = filter(imp_data1, LFSTATP == 1 | LFSTATP == 2)
imp_data2 = filter(imp_data2, LFSTATP == 1 | LFSTATP == 2)
imp_data3 = filter(imp_data3, LFSTATP == 1 | LFSTATP == 2)
imp_data4 = filter(imp_data4, LFSTATP == 1 | LFSTATP == 2)
imp_data5 = filter(imp_data5, LFSTATP == 1 | LFSTATP == 2)

#Filtering for program 9: Health and other related fields
imp_data1 = filter(imp_data1, PGMCIPAP == 9)
imp_data2 = filter(imp_data2, PGMCIPAP == 9)
imp_data3 = filter(imp_data3, PGMCIPAP == 9)
imp_data4 = filter(imp_data4, PGMCIPAP == 9)
imp_data5 = filter(imp_data5, PGMCIPAP == 9)

#Change into binary variables where 1 is employed and 0 is unemployed
imp_data1$LFSTATP = ifelse(imp_data1$LFSTATP == 2,0,1)
imp_data2$LFSTATP = ifelse(imp_data2$LFSTATP == 2,0,1)
imp_data3$LFSTATP = ifelse(imp_data3$LFSTATP == 2,0,1)
imp_data4$LFSTATP = ifelse(imp_data4$LFSTATP == 2,0,1)
imp_data5$LFSTATP = ifelse(imp_data5$LFSTATP == 2,0,1)

#Most relevant features based on LCA analysis (plus LFSTATP)
glm_variables <- c("PGM_P100", "PGM_350", "EDU_010", "STULOANS", "BEF_160", "PAR1GRD", "PAR2GRD", "LFSTATP") 

#convert predictors into factors if it is not yet a factor
imp_data1[glm_variables] = lapply(imp_data1[glm_variables], as.factor)
imp_data2[glm_variables] = lapply(imp_data2[glm_variables], as.factor)
imp_data3[glm_variables] = lapply(imp_data3[glm_variables], as.factor)
imp_data4[glm_variables] = lapply(imp_data4[glm_variables], as.factor)
imp_data5[glm_variables] = lapply(imp_data5[glm_variables], as.factor)

```



Creating training (70%) and test (30%) sets
```{r}
#Creating training and test sets

set.seed(1)
train <- sample(nrow(imp_data1), 0.7*nrow(imp_data1), replace = FALSE)

train_1 <- imp_data1[train,]
train_2 <- imp_data2[train,]
train_3 <- imp_data3[train,]
train_4 <- imp_data4[train,]
train_5 <- imp_data5[train,]


test_1 <- imp_data1[-train,]
test_2 <- imp_data2[-train,]
test_3 <- imp_data3[-train,]
test_4 <- imp_data4[-train,]
test_5 <- imp_data5[-train,]


```

#Logistic regression using glm()

This is to assist in checking the logistic regression assumptions
```{r}

mod_1 <- glm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = imp_data1, subset = train, family = "binomial")
mod_2 <- glm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = imp_data2, subset = train, family = "binomial")
mod_3 <- glm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = imp_data3, subset = train, family = "binomial")
mod_4 <- glm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = imp_data4, subset = train, family = "binomial")
mod_5 <- glm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = imp_data5, subset = train, family = "binomial")

```

#Checking assumptions for logistic regression
```{r}
#Assumption 1: No severe multicollinearity
vif(mod_1)
vif(mod_2)
vif(mod_3)
vif(mod_4)
vif(mod_5)
#Results: Since all the adjusted GVIF values are relatively small (between 1-3), it is a good sign there is very low multicollinearity

#Assumption 2: Checking for influential points using Cook's Distance; Common threshold to compare against is 4/n
cooks_1 <- which(cooks.distance(mod_1) > 4/nrow(imp_data1[train,]))
cooks_1
cooks_2 <- which(cooks.distance(mod_2) > 4/nrow(imp_data2[train,]))
cooks_2
cooks_3 <- which(cooks.distance(mod_3) > 4/nrow(imp_data3[train,]))
cooks_3
cooks_4 <- which(cooks.distance(mod_4) > 4/nrow(imp_data4[train,]))
cooks_4
cooks_5 <- which(cooks.distance(mod_5) > 4/nrow(imp_data5[train,]))
cooks_5


#Results: There are some influential points for each model. However, upon investigation, we realized that they are considered outliers as they have LFSTATP == 0. As a result, we can proceed with caution. 

#Assumption 3: Sample size is large enough

#Assumption 4: Binary outcome variable 

#Assumption 5: Independent observations 


#Reference: Bobbitt, Z. (2020, October 13). The six assumptions of logistic regression (with examples). Statology. https://www.statology.org/assumptions-of-logistic-regression/ 
#Reference: Bobbitt, Z. (2020, November 30). How to Identify Influential Data Points Using Cook’s Distance. Statology. https://www.statology.org/how-to-identify-influential-data-points-using-cooks-distance/
#Reference: Consulted ChatGPT to verify assumptions of logistic regression. Model and implementation were authored by me. OpenAI. (2026). ChatGPT [Large language model]. https://chat.openai.com


```


#Integration of weights
Applying survey weights and creating new model after accounting for them
```{r}

#Creating survey design which will be applied to the svyglm() below for each logistic regression model; ids =~1 assumes independent observations (ie no clustering variables to use)
svy1 <- svydesign(ids = ~1, weights = ~WTPF, data = imp_data1)
svy2 <- svydesign(ids = ~1, weights = ~WTPF, data = imp_data2)
svy3 <- svydesign(ids = ~1, weights = ~WTPF, data = imp_data3)
svy4 <- svydesign(ids = ~1, weights = ~WTPF, data = imp_data4)
svy5 <- svydesign(ids = ~1, weights = ~WTPF, data = imp_data5)

#Creating the logistic regression model and accounting for the weights; quasibinomial() must be used for the family when dealing with binomial output
mod_1 <- svyglm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = train1, design = svy1, family = quasibinomial())
mod_2 <- svyglm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = train2, design = svy2, family = quasibinomial())
mod_3 <- svyglm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = train3, design = svy3, family = quasibinomial())
mod_4 <- svyglm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = train4, design = svy4, family = quasibinomial())
mod_5 <- svyglm(LFSTATP ~ PGM_P100 + PGM_350 + EDU_010 + STULOANS + BEF_160 + PAR1GRD + PAR2GRD, data = train5, design = svy5, family = quasibinomial())

#Reference: Geeks for Geeks. (2025, April 28). Survey package in r. Geeks for Geeks. https://www.geeksforgeeks.org/r-language/survey-package-in-r/
#Reference: Consulted ChatGPT to better understand how to apply survey weights to logistic regression. Model and implementation were authored by me. OpenAI. (2026). ChatGPT [Large language model]. https://chat.openai.com

```


#Pooling results 
```{r}

#Pooling the results of all five logistic regression model after accounting for the weights
final_pooled <- pool(list(mod_1, mod_2, mod_3, mod_4, mod_5))
sum_pooled <- summary(final_pooled)
sum_pooled

#Odds ratio of intercept since it has a pvalue smaller than 0.05
exp(sum_pooled[1,2])

#Setting up summary dataframe for predictions in r chunks below
model_number <- c(1:5)
summary <- data.frame(model_number)

#Reference: Consulted ChatGPT to learn how to pool results through Rubin's rule. Model and implementation were authored by me. OpenAI. (2026). ChatGPT [Large language model]. https://chat.openai.com


```

#Predicting employment outcome based on field of study

There are five models because each model corresponds to a different test set since there are five imputed datasets

  
Model 1
```{r}
pred_1 <- predict(mod_1, test_1, type = "response") #predicting probabilities using test set

actual = as.numeric(as.character(test_1$LFSTATP)) #converting actual classes to numeric

# Create ROC curve
roc_curve <- roc(actual, pred_1)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve")


#find best threshold
coordinate_thresh <- pROC::coords(roc_curve,"best", ret = "threshold")
optimal_threshold <- coordinate_thresh$threshold


pred_data = rep(0,nrow(test_1)) #Assign classes 
pred_data[pred_1 > optimal_threshold] = 1 #Assign classes based on optimal threshold
summary[1,2] = mean(actual==pred_data) #Prediction accuracy
conf.mat <- table(actual, pred_data) #Confusion Matrix 

#calculating precision
TP <- conf.mat[2,2]
FP <- conf.mat[1,2]
summary[1,3] <-  TP / (TP + FP)

#calculating recall
FN <- conf.mat[2,1]
summary[1,4] <-  TP / (TP + FN)

# AUC 
summary[1,5] <- pROC::auc(roc_curve)

#f1 
summary[1,6] <- 2*((summary[1,3]*summary[1,4])/(summary[1,3]+summary[1,4]))

#Reference: Geeks for Geeks. (2025, July 23). Deciding threshold for glm logistic regression model in R. Geeks for Geeks. https://www.geeksforgeeks.org/machine-learning/deciding-threshold-for-glm-logistic-regression-model-in-r/ 
#Reference:Geeks for Geeks. (2025, April 9). Precision, recall, and f-1 score using r. Geeks for Geeks. https://www.geeksforgeeks.org/r-language/precision-recall-and-f1-score-using-r/ 
  
```


Model 2
```{r}

pred_2 <- predict(mod_2, test_2, type = "response") #predicting probabilities 

actual = as.numeric(as.character(test_2$LFSTATP)) #converting actual class into numeric

# Create ROC curve
roc_curve <- roc(actual, pred_2)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve")

#find best threshold
coordinate_thresh <- pROC::coords(roc_curve,"best", ret = "threshold")
optimal_threshold <- coordinate_thresh$threshold


pred_data = rep(0,nrow(test_2)) #Assign classes 
pred_data[pred_2 > optimal_threshold] = 1 #Assign classes based on optimal threshold
summary[2,2] = mean(actual==pred_data) #Prediction accuracy

conf.mat <- table(actual, pred_data) #Confusion Matrix 


#calculating precision
TP <- conf.mat[2,2]
FP <- conf.mat[1,2]
summary[2,3]<-  TP / (TP + FP)

#calculating recall
FN <- conf.mat[2,1]
summary[2,4] <-  TP / (TP + FN)

# AUC 
summary[2,5] <- pROC::auc(roc_curve)

#f1 
summary[2,6] <- 2*((summary[2,3]*summary[2,4])/(summary[2,3]+summary[2,4]))

```


Model 3
```{r}


pred_3 <- predict(mod_3, test_3, type = "response") #predicting probabilities 

actual = as.numeric(as.character(test_3$LFSTATP)) #converting actual class into numeric

# Create ROC curve
roc_curve <- roc(actual, pred_3)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve")

#find best threshold
coordinate_thresh <- pROC::coords(roc_curve,"best", ret = "threshold")
optimal_threshold <- coordinate_thresh$threshold

pred_data = rep(0,nrow(test_3)) #Assign classes 
pred_data[pred_3 > optimal_threshold] = 1 #Assign classes based on optimal threshold
summary[3,2] = mean(actual==pred_data) #Prediction accuracy

conf.mat <- table(actual, pred_data) #Confusion Matrix 


#calculating precision
TP <- conf.mat[2,2]
FP <- conf.mat[1,2]
summary[3,3] <-  TP / (TP + FP)

#calculating recall
FN <- conf.mat[2,1]
summary[3,4] <-  TP / (TP + FN)

# AUC 
summary[3,5] <- pROC::auc(roc_curve)

#f1 
summary[3,6] <- 2*((summary[3,3]*summary[3,4])/(summary[3,3]+summary[3,4]))

```


Model 4
```{r}

pred_4 <- predict(mod_4, test_4, type = "response") #predicting probabilities 

actual = as.numeric(as.character(test_4$LFSTATP)) #converting actual class into numeric

# Create ROC curve
roc_curve <- roc(actual, pred_4)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve")

#find best threshold
coordinate_thresh <- pROC::coords(roc_curve,"best", ret = "threshold")
optimal_threshold <- coordinate_thresh$threshold


pred_data = rep(0,nrow(test_4)) #Assign classes 
pred_data[pred_4 > optimal_threshold] = 1 #Assign classes based on optimal threshold
summary[4,2] = mean(actual==pred_data) #Prediction accuracy
conf.mat <- table(actual, pred_data) #Confusion Matrix 

#calculating precision
TP <- conf.mat[2,2]
FP <- conf.mat[1,2]
summary[4,3] <-  TP / (TP + FP)

#calculating recall
FN <- conf.mat[2,1]
summary[4,4] <-  TP / (TP + FN)

# AUC 
summary[4,5] <- pROC::auc(roc_curve)

#f1 
summary[4,6] <- 2*((summary[4,3]*summary[4,4])/(summary[4,3]+summary[4,4]))


```


Model 5
```{r}

pred_5 <- predict(mod_5, test_5, type = "response")

actual = as.numeric(as.character(test_5$LFSTATP)) #converting actual class into numeric

# Create ROC curve
roc_curve <- roc(actual, pred_5)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve")

#find best threshold
coordinate_thresh <- pROC::coords(roc_curve,"best", ret = "threshold")
optimal_threshold <- coordinate_thresh$threshold

pred_data = rep(0,nrow(test_5)) #Assign classes 
pred_data[pred_5 >optimal_threshold] = 1 #Assign classes based on optimal threshold
summary[5,2] = mean(actual==pred_data) #Prediction accuracy
conf.mat <- table(actual, pred_data) #Confusion Matrix 

#calculating precision
TP <- conf.mat[2,2]
FP <- conf.mat[1,2]
summary[5,3] <-  TP / (TP + FP)

#calculating recall
FN <- conf.mat[2,1]
summary[5,4] <-  TP / (TP + FN)

# AUC 
summary[5,5] <- pROC::auc(roc_curve)

#f1 
summary[5,6] <- 2*((summary[5,3]*summary[5,4])/(summary[5,3]+summary[5,4]))

```




#Summary of Performance Metrics
Summary of all metrics per model
```{r}
colnames(summary) <-c("Model", "Accuracy", "Precision", "Recall", "AUC", "F1")
View(summary)

mean_all <- sapply(summary[,2:6], mean)
mean_all #average performance metrics across all five models


```